\input{header.tex}

\title{Overview}
\begin{document}
\maketitle


This is an experiment. I need your feedback to make sure it's working
for you.


\section{Background}


Have you taken

\begin{enumerate}
\item A linear algebra course
\item A differential equations course
\item A course on signals and systems
\item A DSP course
\item A stochastic process theory course
\item Have you called fft in the past year
\end{enumerate}

\section{Introduction}
The importance of consistent pedagogy -- $x(t)$ vs $x[n]$ 

The unimportance of formalism -- this is a rapid survey (drinking from the 
firehose) and the formalism sometimes gets in the way. That said, if you
wish to do actual work in this area, it's important to at least collaborate
with someone who knows it. 

\section{What we don't cover}
\begin{itemize}

\item \textbf{Control theory}: Control theory is one of
  the great successes of 20th century engineering, and has enabled so
  much of the amazing devices we have today. Laurent Lessard put
  together a great introduction that I can refer you to if you're
  interested. There are need connections between control theory and
  recent active learning work (which we will not talk about) 
\item \textbf{Stochastic processes}: The real world is stochastic, as is most of your
  data. Stochastic process theory formalizes this, and is something
  even CS students should know, as it underlies large swaths of
  machine learning. That said, LTI systems like those we discuss
  impact simplistic stochastic processes in about the way you would
  expect, so there's not a lot of ``surprising'' material here.
\item \textbf{Filter design}. Of course, once you have this incredible
  framework for signal processing, you want to then process some
  signals, and to do this you often ened to engineer filters. This
  used to be a very deep topic that has been mostly reduced to a
  series of matlab function calls.
\item \textbf{ advanced modulation schemes / information theory} The
  bleeding edge of the PHY layer involves state-of-the-art modulation
  schemes as well as lots of recent work in coding theory. We skip all
  of this .
\end{itemize} 

\section{Format}
\begin{itemize}
\item 7 lectures
\item heavy on background at first
\item Lots of neat examples (i hope)
\item If you want supplemental material let me know
\end{itemize} 

\section{What I hope to achieve}
\begin{enumerate}
\item Get people excited about these problems
\item in two years you remember enough keywords to intelligently
  google the right thing and not have the wikipedia page scare you
\item Realize that computer science has been doing a lot of this stuff
  forever, and that you really do have the background and skills to
  contribute.
\end{enumerate} 

\end{document}
